### 2019-12-04 results/sec_2-work_load.tex
Added Database Structure details

\section{Work Load Creation}

Precipitation
\begin{lstlisting}[language=sh, caption=Test Plan Help]
-h | --help: Usage
    <ROOT_DIR> <COMMAND> <REQ_SIZE>
    - REQ_SIZE (optional): 24(1) | 288(2) | 1044(3)
  NOTE: Modify test.conf as necessary
  e.g.
  test_plan.sh ~/wdias/wdias-performance-test run
  test_plan.sh ~/wdias/wdias-performance-test run 2
  - Run all the steps in order of setup, import, 
    create_extension, extension, export, all, query

  test_plan.sh ~/wdias/wdias-performance-test setup
  test_plan.sh ~/wdias/wdias-performance-test import 24
  test_plan.sh ~/wdias/wdias-performance-test create_extension 24
  test_plan.sh ~/wdias/wdias-performance-test extension 24
  test_plan.sh ~/wdias/wdias-performance-test export 24
  test_plan.sh ~/wdias/wdias-performance-test all 24
  test_plan.sh ~/wdias/wdias-performance-test query 24

  Distributed Mode
  SERVER_IPS=<IP1,IP2...> test_plan.sh ~/wdias/wdias-performance-test run

\end{lstlisting}

Water level