### 2019-12-04 results/sec_4-conclusion.tex

\section{Conclusion}

Throughput

Latency

Resource Utilization

Auto Scaling/Scalability

Risk of unable to process data