### 2020-02-06 summary/sec_2-future_work.tex
Added All 1440 result graphs

\section{Future Work}
- Lack of data prepossessing modules
- Supporting Irregular Grids
- Grid performance
- Define infrastructure as code - Teraform
- Improve test cases - correct RPS
- Further improve hierarchical database performance
- Add Alerting extension modules
- Add pubsub extension modules