### 2020-02-08 summary/sec_1-summary.tex
Updated future works

\section{Summary}

As summary of the proposed system, it's performance and conclusion over the results.