### 2020-02-09 introduction/sec_2-problem_statement.tex
thesis - Move introduction into multiple files

\section{Problem Statement}
\dbc{Convert following bullets to a paragraph. Goals belongs to next section. What we need here is a set of sentences highlighting the needs that are used to derive goals. Then provide a single sentence problem statement. Look at some of the thesis I shared via OneNote and my website for example problem statements.}
\begin{itemize}
    \item Handling spatio-temporal weather data is challenging
    \item Need an extendable system to support weather models by efficiently storing and retrieving data
    \item Goal is to design and develop a Weather Data Integration and Assimilation System that,
    \begin{itemize}
        \item Integrates weather data from different sources with quality control
        \item Efficiently store spatio-temporal weather data with multiple dimensions
        \item Support geographical and time-base queries
    \end{itemize}
\end{itemize}