### 2020-02-11 introduction/sec_2-problem_and_objectives.tex

\section{Problem Statement}
Handling spatio-temporal weather data is challenging \db{as ...}. Moreover, the data integration system needs to be capable of interacting with different data formats including spatio-temporal bulk stream data. Furthermore, it should support geographical and time-base queries, as well as provide third-party access to data with easy integration. Scalability and high availability are also essential to support different use cases and mission-critical nature of applications. Therefore, we need an extendable system to support weather models by efficiently ingesting, storing, and retrieving data. In this context, the research problem can be stated as follows:

\emph{....}
\dbc{Give a proper problem statement.}

\section{Objectives}
We address above problem statement by achieving the following research objectives:
\begin{itemize}
    \item To develop an open platform to integrate weather data from different sources in different formats
    \item To design and develop a schema to store multidimensional weather timeseries data while optimizing access time, availability, and storage
    \item To optimize schema to support data queries based on geography and time
    \item To develop an API for third-party data sharing
%    \item To use open source tools and provide the source publicly with appropriate liecense.
\end{itemize}