### 2020-02-11 results/sec_0-introduction.tex
Reviewed again after PGS reject

%\section{Introduction}

This chapter includes performance study of the \acrshort{wdias}. In Section \ref{se:test_plan} describes the planned method of testing the system.
In the next section \ref{se:workload} describes the logic behind creating the workload to do the performance testing planned in \ref{se:test_plan}.
Following section \ref{se:observations} includes the results observed while performance study. And summarised the observations in the next section \ref{se:discussion}.