### 2020-02-13 results/sec_0-introduction.tex
Fixed: “Section”
- Add “Section”
- “section” -> “Section”

%\section{Introduction}

This chapter includes performance study of the \acrshort{wdias}. In Section \ref{se:test_plan} describes the planned method of testing the system.
In the next Section \ref{se:workload} describes the logic behind creating the workload to do the performance testing planned in Section \ref{se:test_plan}.
Following Section \ref{se:observations} includes the results observed while performance study. And summarised the observations in the next Section \ref{se:discussion}.