### 2020-02-20 introduction/sec_3-outline.tex

\section{Outline}
The rest of the thesis is organized as follows. The literature review is presented in \cref{ch:literature}, and discuss the existing Weather Data Integration Systems such as \acrshort{lead}, \acrshort{fews}, \acrshort{madis}, and \acrshort{dias}. Solution approach with the proposed system architectures and adaptation with problem occurred are presented in Chapter \ref{ch:method}. It provides details about how the \acrshort{wdias} is designed by following a microservice architecture and the data preprocessing capabilities of the system. Performance analysis is presented in Chapter \ref{ch:results}. Here we discuss the experimental setup of the \acrshort{wdias} and analyze the observations concerning performance metrics. Finally, in Chapter \ref{ch:summary}, we conclude and summarize the work and further discussed the research limitations and future work.