### 2020-02-21 method/sec_0_introduction.tex
Plagiarism microservices

This chapter includes the design decision taken while implementing the \acrfull{wdias}, and how is the system got adapted from the systems analyzed in Chapter \ref{ch:literature}.
\cref{se:high_level_design} includes a brief discussion on how the weather forecasting is happening and why it needs to interact with multiple data formats is presented.
\cref{se:architectural_decisions} discussed about how did we evolve over the architectural design while designing and implementing the \acrshort{wdias}.
\cref{se:microservice} gives a better understanding of the microservice architectural concepts that follow while implementing the \acrshort{wdias}.
Then the reasons behind developing \acrshort{wdias} database structure and technologies use for that discussed in \cref{se:db_struct}.
\cref{se:data_preprocess} discussed how the \acrshort{wdias} enables for data preprocessing and adding capabilities via the concept of Extensions.
Final \cref{se:query} includes details on how to perform timeseries search and Geo-based queries on the \acrshort{wdias}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{High-level Design}
\label{se:high_level_design}

To get an overall idea of the flow of weather prediction, let us look into one of the flows using at \acrshort{curw} \cite{CUrWSL2017SL}.
\dbc{Use CUrW-SL}
\gkc{FIXED}

\begin{enumerate}
    \item Extract \acrshort{GRIB} \acrshort{wrf} model data to \acrshort{netCDF}
    \item From \acrshort{netCDF} to \acrshort{csv} of each location data, then feed to HEC-HMS hydro-dynamic model and get the output as \acrshort{csv}
    \item \acrshort{csv} and Rain cell Grid feed into the FLO2D hydro-logic model, and get the water level as ASCII Grid files after completion of simulation
    \item Upload ASCII files to ArcGIS create Raster files, then use for the LOSS Estimation
\end{enumerate}
When looked into the above flow, it gives a clear understanding of multiple data formats using while weather predictions.
Other than that, to correct the model outputs, it required to collect real-time data such as;
\begin{itemize}
    \item Automated weather station sending data in higher frequent intervals via HTTP protocol
    \item Automated water level stations send data in higher frequency
\end{itemize}
Other than above many data formats using for handling data based on devices and standards which are using. Also, for data modeling, it needs to convert the data into model compatible data format, then after successfully run the model it needs to collect data from the model output data formats.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Modules of a Weather Data System}
\label{subse:modules_weather_data_integration_sys}

\cref{fi:wdia_components} shows the basic functions of a Weather data integration and assimilation system which consists of the following modules:

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.8\textwidth]{method/misc/weather_data_system_components.jpg}}
\caption{Modules of a weather data system.}
\label{fi:wdia_components}
\end{figure}

\subsubsection{Integration}
The system should capable of integrating data from different sources such as satellite data, high end, and low-end weather station, etc. And the system should be able to handle multidimensional spatial and temporal weather data efficiently and optimally. 
\subsubsection{Assimilation}
Then the system should be able to fulfill weather models varying data requirements. Also, those models reproduce a large set of redundant data, thus the system should store the bulk data while optimizing the disk space.
\subsubsection{Dissemination}
Then different users should be able to retrieve data as they required. Also, the users should be able to easy to search into the available that in the system base on timeseries metadata or based on Geo queries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \paragraph{Timeseries}-- A timeseries is simply a series of data points ordered in time. In weather domain, it interest in timeseries in perspective of observations to forecasting. Each timeseries, the data points can be form in different formats as well. For example, scalar (0D), vector (1D), grid (2D), and polygon (2D).

\dbc{Why are you referring to timeseries here? It's not one of the 3 key components in Fig. 3.1.}
\gkc{Move this to WDIAS microservices \cref{sebse:wdias_microservices}. It need to describe the reason behind decompose into microservices. Seems to be this is confusing at this point.}

% In the proof of concept design described below, WDIAS focuses on handling scalar, vector and grid timeseries data. However, the system could be extended to handle polygon-based timeseries as well.

% \subsubsection{Base of WDIAS architecture}
\dbc{What's Base of WDIAS architecture?}
\gkc{Moved to WDIAS database structure section \cref{sebse:wdias_microservices}}

% \hl{The base of WDIAS architecture origins based on attributes of a timeseries.} There are many attributes to differentiate timeseries one from another. But 
% Among many attribute to differentiate a timeseries from another, the following can be consider as the key attributes in uniquely identify a timeseries from another: % for the \acrshort{wdias}.
\dbc{Highlighted sentence doesn't make sense to me. All yellow highlights mean they are not clear.}

% \begin{itemize}
%     \item Module -- module that generates the timeseries
%     \item Value type -- Scalar, vector, grid, or polygon
%     \item Location of readings
%     \item Parameter -- measuared parameter
%     \item hl{Timeseries type}
%     \begin{itemize}
%         \item Sources -- External or simulated
%         \item Category -- Historical or forecast
%     \end{itemize}
%     \item Time step
% \end{itemize}
% \dbc{Use bullets only when essential. Other places use paragraphs.}

% A detail description of the above attributes are presented in \cref{se:db_struct}.