### 2020-04-25 results/sec_0-introduction.tex
Up to results->workload

%\section{Introduction}

This chapter includes the performance study of the \acrshort{wdias}. \cref{se:test_plan} describes the test plans for the performance testing.
\cref{se:workload} describes the logic behind creating the workload to do the performance testing planned in \cref{se:test_plan}.
\cref{se:observations} includes the results observed during performance study and summarised the observations in \cref{se:discussion}.