### 2020-04-26 glossary.tex
Updated -> Method - Microservices

\newacronym{wdias}{WDIAS}{Weather Data Integration and Assimilation System}

\newacronym{fews}{Delft-FEWS} {Deltares FEWS}
\newacronym{lead}{LEAD}{Linked Environments for Atmospheric Discovery}
\newacronym{dias}{DIAS}{Data Integration and Assimilation System}
\newacronym{madis}{MADIS}{Meteorological Assimilation Data Ingest System}

\newacronym{nwm}{NWMs}{Numerical Weather Models}
\newacronym{soa}{SOA}{Service Oriented Architecture}
\newacronym{wrf}{WRF}{Weather Research and Forecast}
\newacronym{esb}{ESB}{Enterprise Service Bus}
\newacronym{microservice}{Microservice}{Microservice Architecture}

\newacronym{netCDF}{netCDF}{Network Common Data Form}
\newacronym{GRIB}{GRIB}{General Regularly-distributed Information in Binary form}
\newacronym{csv}{CSV}{Comma-separated Values}

\newacronym{rdbms}{RDBMS}{Relational Database Management System}
\newacronym{k8s}{K8s}{Kubernetes}
\newacronym{eks}{Amazon EKS}{Amazon Elastic Kubernetes Service}

\newacronym{go}{Go Lang}{Go Programming Language}
\newacronym{rps}{RPS}{Requests Per Second}
\newacronym{api}{API}{Application Programming Interface}
\newacronym{rest}{REST}{Representational State Transfer}

\newacronym{curw}{CUrW-SL}{Center for Urban Water, Sri Lanka}