### 2020-05-31 method/sec_6-query.tex
Proofread Chapter 4

\section{Query Timeseries}
\label{se:query}

We used the query adapter primarily for supporting geo-spatial queries. As the primary service for searching for timeseries, it supports retrieval of timeseries metadata for external queries as well. Appendix B described the search query capabilities within \acrshort{wdias}.

The system provides timeseries metadata query endpoints to search for existing timeseries metadata within the system using primary key attributes.
Also, it provides a set of geo-query endpoints, and users can perform geo-search queries with sending geo-information in geoJson format. The following geo-queries are supported by the system, and a detailed description is presented in Appendix B.
\begin{itemize}
    \item Query locations within area
    \item Query parameters in locations
    \item Query timeseries by location
    \item Query timeseries by locations
    \item Query timeseries by Parameter
    \item Query timeseries in area
    \item Query timeseries in area by parameter
    \item Query all timeseries
\end{itemize}