### 2020-12-25 introduction/sec_3-outline.tex

\section{Outline}
The organization of the rest of the thesis is as follows; In \cref{ch:literature}, the literature review is presented. It discusses the existing weather data integration systems such as \acrshort{lead}, \acrshort{fews}, \acrshort{madis}, \acrshort{dias}, and \acrshort{msm}. Solution approach including the proposed system architectures and adaptation with problem occurred are presented in Chapter \ref{ch:method}. It provides details on how the \acrshort{wdias} is designed by following a microservice architecture and its data pre-processing capabilities. Performance analysis is presented in Chapter \ref{ch:results}. Here we discuss the experimental setup of the \acrshort{wdias} and analyze the observations concerning performance metrics. Finally, Chapter \ref{ch:summary}, summarizes the research work and further discusses the research limitations and future work.