### 2020-12-30 method/sec_6-query.tex

\section{Query Timeseries}
\label{se:query}

We used the query adapter primarily for supporting geo-spatial queries. As the primary service for searching for timeseries, it supports retrieval of timeseries metadata for external queries as well. Appendix B described the search query capabilities within \acrshort{wdias}.

The system provides timeseries metadata query endpoints to search for existing timeseries metadata within the system using primary key attributes mentioned in the \cref{subse:timeseries_key_attributes}.
Also, it provides a set of geo-query endpoints, and users can perform geo-search queries with sending geo-information in geoJson format. The following geo-queries are supported by the system, and a detailed description is presented in Appendix B.
\begin{itemize}
    \item Search for available locations within the provided geoJson area in the query.
    \item Search for available parameters in provided locations in the query.
    \item Search for available timeseries for a provided location in the query.
    \item Search for available timeseries for the given locations in the query.
    \item Search for available timeseries for given parameter. E.g., get available waterlevel timeseries in the system.
    \item Search for available timeseries within the provided geoJson area in the query.
    \item Search for available timeseries within the provided geoJson area by parameter. E.g., get available waterlevel timeseries within a given area.
    \item Query for all timeseries in the system.
\end{itemize}